---
title: "Socio-Cultural Influences on Depressive Disorders in the US: An Analysis from 2020 to 2024"
subtitle: "Exploring the Impact of Age, Gender, Race, and Education on Mental Health Trends During and After the COVID-19 Pandemic"
author: Chay Park
thanks: "Code and data are available at: https://github.com/Chay-HyunminPark/Anxiety_depressive_disorder."
date: today
date-format: long
abstract: "This study explores how socio-cultural factors like age, sex, race/ethnicity, and education level influenced depressive disorder in the U.S. during and after the COVID-19 pandemic, analyzing data from 2020 to 2024. It uncovered variations in mental health impacts across different demographic groups. The research reveals a complex relationship between socio-cultural factors and mental health, highlighting the increased vulnerability of certain populations during global crises. The findings contribute to the development of targeted public health measures to alleviate mental health issues amid and following pandemics, enhancing our understanding of the impact of socio-cultural factors on mental well-being in times of crisis."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# Load necessary libraries for the paper
library(tidyverse)
library(lubridate)
library(arrow)
library(here)
library(ggrepel)  # For better label placement
library(extrafont)
library(ggplot2)
library(gridExtra)

```


# Introduction

Depression and anxiety are not only prevalent but escalating issues in the modern world, affecting individuals across the globe regardless of demographic factors (World Health Organization, 2021). These mental health challenges have been further exacerbated by the COVID-19 pandemic, which led to widespread social and economic upheaval (Brooks et al., 2020). While various studies have explored the impact of the pandemic on mental health, there remains a significant gap in understanding the interactions between socio-cultural factors and the prevalence of anxiety and depressive disorders during this period (Pfefferbaum & North, 2020).

This paper addresses this gap by analyzing data from the U.S. Census Bureau's Household Pulse Survey, initiated in response to the pandemic to assess its effects on American households, including mental wellness (U.S. Census Bureau, 2020). Our analysis focuses on the correlation between the prevalence of anxiety or depressive disorders and socio-cultural factors such as age, sex, race/ethnicity, and education level during the pandemic and post pandemic years of 2020 to 2024. Specifically, the estimand of this study is the difference in the rates of symptom of depressive disorders across individuals, correlated with socio-cultural factors including age, sex, race/ethnicity, and education level.

Employing a quantitative methodology, this paper utilizes R (R Core Team 2020) for analysis. We utilized packages like ggplot2 (Wickham et al., 2016) and gridExtra (Murrell, 2021) for data analysis and visualization. Our findings reveal significant trends and disparities in the rates of anxiety and depressive disorders across different socio-cultural groups, highlighting the intricate relationship between these factors and mental health outcomes during the pandemic.

The importance of this research lies in its potential to inform public health policies and interventions, aiming to mitigate the adverse effects of such global crises on mental health (Xiong et al., 2020). By understanding the specific socio-cultural factors that contribute to increased vulnerability to anxiety and depressive disorders, targeted strategies can be developed to support affected populations more effectively.

The paper is structured to first present the trend analysis of anxiety and depressive disorder rates in the U.S. from 2020 to 2024, followed by an examination of how these rates vary across different demographic groups. Subsequent sections discuss the statistical methods and results for each research question. It opens up the findings and let the world explores their broader implications for mental health research and policy. The conclusion summarizes the key insights, acknowledges the limitations of the study, and suggests avenues for future research.

The study focuses on answering the following research questions:

* What is the trend in US anxiety and depressive disorder rates from 2020 to 2024?

* How do depressive rates vary across different age groups?

* How do depressive rates differ based on race and ethnicity?

* How do depressive rates differ based on sex?

* How do depressive rates differ based on education level?

Through this analysis, we aim to better understand contributing factors of mental wellness of the US population. Furthermore, we display possible correlation between factors and the mental wellness states. 

# Data {#sec-data}

## Methodology

The U.S. Census Bureau, in collaboration with five federal agencies, launched the Household Pulse Survey to produce data on the social and economic impacts of Covid-19 on American households. The Household Pulse Survey was designed to gauge the impact of the pandemic on employment status, consumer spending, food security, housing, education disruptions, and dimensions of physical and mental wellness (National Center for Health Statistics, 2024). In the paper, we will focus on the impact of the pandemic on mental wellness by analysing the dataset of the indicators of anxiety or depression based on reported frequency of symptoms. The data on the trend in US anxiety and depressive disorder rates, measured from 2020 to 2024, was sourced from the Centers for Disease Control and Prevention (CDC). Alongside with 'National Estimate' group in which covers the whole population trends, trends in anxiety and depressive disorder rates by population subgroup, specifically categorized by age, race and ethnicity, sex, education groups were gathered from the National Center for Health Statistics (NCHS). The data is freely available at https://data.cdc.gov/NCHS/Indicators-of-Anxiety-or-Depression-Based-on-Repor/8pt5-q6wp/about_data, with raw files located in the inputs/data on the GitHub repository. 

## Features

In this paper, data validation is conducted using an R script file named '03-test_data' to ensure the accuracy of the dataset obtained directly from the source. The paper utilizes both 'Indicators_of_Anxiety_or_Depression_Based_on_Reported_Frequency_of_Symptoms_During_
Last_7_Days_20240401' and 'Lack_of_Social_Connection' CSV files. Then by the characteristics of the group, the data sets are separately saved as 'Age_subgroup_trends' where it only saves the rows with 'By Age' under the 'Group' column, irrespective of race and ethnicity. Similarly, 'Educ_subgroup_trends' are the dataset only with the rows 'By Education' under the 'Group' column to identify the patterns of each population subgroup with classification of highest educational attainment, regardless of race and ethnicity. The column 'Value' is used to denote the symptom rate on the y-axis. Column 'Time Period End Date' is used to denote the range of the period, a date for the x-axis.


# Results

## Trend in US Anxiety and Depressive Disorder Rates

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-andtrend-graph
#| fig-cap: "Trend in US Anxiety and Depressive Disorder Rates"

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Overall_trends.parquet"))

# Filter the dataset for the specified symptoms
filtered_data <- data %>%
  filter(Indicator %in% c('Symptoms of Depressive Disorder', 
                          'Symptoms of Anxiety Disorder', 
                          'Symptoms of Anxiety Disorder or Depressive Disorder')) %>%
  mutate(Indicator = factor(Indicator, levels = c('Symptoms of Anxiety Disorder or Depressive Disorder',
                                                  'Symptoms of Anxiety Disorder',
                                                  'Symptoms of Depressive Disorder')))

# Ensure that 'Time.Period.End.Date' is a Date object
filtered_data$Date <- as.Date(filtered_data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Plotting the line graph
p <- ggplot(filtered_data, aes(x = Date, y = Value, group = Indicator, color = Indicator)) +
  geom_line() +
  geom_hline(yintercept = seq(10, 45, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL,
       y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.direction = "vertical",
        legend.text = element_text(family = "Times"),
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

The overall trend in US anxiety and depressive disorder rates @fig-andtrend-graph shows significant fluctuations over the observed period, with several notable peaks indicating heightened rates of reported symptoms. For instance, one of the highest peaks occurs in early 2021, where the rate of symptoms for combined anxiety and depressive disorders reaches approximately 45%. This peak reflects a substantial increase compared to earlier peeriods. Another significant peak, observed in mid-2022, sees rates approaching closer to 36-37%, illustrating another period of notable increases in symptoms among the population. 

These peaks are interspersed with valleys, such as in late 2020, where rates temporarily din to around 25%, demonstrating the episodic nature of these conditions. The plot also highlights a general trend of increasing rates from the onset of the data in 2020 through to peaks in 2021, followed by some fluctutation but maintaining relatively high levels overall. 

In late 2023, there is a notable decline in the symptom rates for both anxiety and depressive disorders, suggesting a potential easing of these mental health challenges or the effectiveness of interventions and strategies implemented over the period. This decrease is significant as it may indicate periods of relief amidst the prevailing trends of increasing disorder rates, which could be crucial for future public health planning and policymaking.

The @fig-andtrend-graph provides a line graph visualization of these rates over time. This visual representation allows for clear observation of trends and seasonal variations, with specific dates marked on the x-axis to aid in pinpointing changes. 

## Trends in Depressive Disorder Rates by Population Subgroup

### Population subgroup by age

Overall, the younger the age group is, the higher the symptom rates are. @fig-ageandtrend-graph illustrates the trends in depressive disorder rates by population subgroup segmented by age, covering the period from May 2020 to March 2024. This graph reveals several key insights about how depressive symptoms have fluctuated across different age groups during and potentially influenced by the ongoing global and societal changes. The line graph distinguishes each age group with a separate line, showing the trend of depressive disorder rates for each demographic:

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-ageandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Age"

# Load necessary libraries
library(tidyverse)
library(arrow)
library(here)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Age_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_hline(yintercept = seq(0, 50, by = 10), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "bottom",  # Display legend at the bottom of the plot
        legend.title = element_blank(),  # Remove the legend title if desired
        legend.box = "horizontal",  # Arrange legend items horizontally
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

1. **Young Adults (18-29 years):** This group shows relatively higher rates of depressive symptoms throughout the period, with peaks that suggest increased vulnerability during certain times. Notable spikes might correlate with specific events or seasons, indicating external impacts on mental health. During the period, the symptom rates of this group are generally 10% higher than the second highest age group. Moreover, it is observed that there is 20% difference gap with the lowest symptom rates age group, 80 years and above. 

2. **Middle-aged Groups (40-59 years):**This line show moderate fluctuation but generally less variance compared to younger adults or older adults. Middle-aged groups generally locate in between 30 -39 years young adult group and 60 - 69 years old adults group. However, on 2023 May, middle-aged group 40 - 49 years exceeds 30 - 39 years young adults group by a few percentage of symptom rates. Overall, both 40 - 49 years age group and 50 - 59 years age group demonstrate similar trends with some points where symptom rates of 50 - 59 years age group surpassed 40 - 49 years age group. For instance, 2021 March and 2022 September. 

3. **Old Adults (80 years and above):** Most fluctuating age groups, with a peak in 2021 January and multiple lowest peaks including 2021 July, 2022 March, and 2022 August. This group still score below than majority of the age groups but there are several moments where it overtakes the symptom rates of 70 - 79 years old group. 

### Population subgroup by sex

@fig-sexandtrend-graph differentiates between the two sex subgroups, highlighting the trends in depressive disorder rates for each:

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-sexandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Sex"

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Sex_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Filter the data to get the rows for the specific date to label
label_date <- as.Date("2022-01-10")
label_data <- data %>%
  filter(Date == label_date)

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",  # Remove the separate legend
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

p + geom_text_repel(
  data = label_data,
  aes(label = Subgroup),
  nudge_x = 0,  
  nudge_y = 0,
  direction = "y",
  color = "black",
  family = "Times"
)

```

1. **Female Subgroup:** This line consistently shows higher rates of depressive symptoms compared to the male subgroup. The trend line for females exhibits several pronounced peaks. Two major peaks happened in 2020 August and 2021 January. And then as of January 2021 the trends starting to decline up until June 2021. From there the trends are following sideways trends with minor peaks and troughs along the way until the end of 2023.  

2. **Male Subgroup:** Although the rates for the male subgroup are generally lower than those for females, the trend line still shows significant fluctuation with similar peaks and troughs, indicating that males are also significantly affected by the same factors that increase depressive symptoms. Reaching the end of 2023, the line graphs for both population subgroups almost align with each other and exhibits almost identical behaviours. 

### Trends in Depressive Disorder Rates by Population Subgroup by Race and Ethnicity

The analysis of depressive disorder rates among various racial and ethnic groups @fig-raceandtrend-graph reveals distinct patterns of mental health disparities from 2020 to 2024. The data, visualized through a line graph, indicates varying symptom rates across five subgroups: Hispanic, Caucasian, African American, Asian, and Other. 

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-raceandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Race"

# Load necessary libraries
library(tidyverse)
library(arrow)
library(here)
library(ggrepel)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Race_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Rename the subgroup values
data$Subgroup <- recode(data$Subgroup,
                        "Hispanic or Latino" = "Hispanic",
                        "Non-Hispanic Asian, single race" = "Asian",
                        "Non-Hispanic Black, single race" = "African American",
                        "Non-Hispanic White, single race" = "Caucasian",
                        "Non-Hispanic, other races and multiple races" = "Other")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Filter the data to get the rows for the specific date to label
label_date <- as.Date("2022-01-10")
label_data <- data %>%
  filter(Date == label_date)

# Plotting the line graph
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_text_repel(data = label_data, aes(label = Subgroup, x = Date, y = Value), 
                  nudge_x = 0, nudge_y = 0, family = "Times",
                  size = 3,  # Make label text smaller
                  arrow = arrow(type = "closed", length = unit(0.02, "inches"), ends = "last", angle = 30)) +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = NULL, y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

# Display the plot
print(p)

```

1. **Hispanic and African American Subgroups:** Both subgroups exhibit relatively higer rates of depressive symptoms throughout the observed period compared to other groups. This may suggest a greater impact of socio-economic and possibly systemic factors affecting these communities disproportionately. The maximum rates observed for Hispanics and African Americans were 41.10% and 41.43%, respectively, with Hispanics showing a minimum rate significantly lower at 18.77%, suggesting more substantial fluctuatinos. 

2. **Caucasian Subgroup:** While the Caucasian group shows lower symptom rates than the Hispanic and African American groups, it follows a similar pattern of fluctuations over time. The maximum rate for Caucasians was 35.17%, with an average rate of 27.92%, indicating that while the intensity may vary, the underlying factors influencing depressive symptoms are somewhat consistent across these groups.

3. **Asian Subgroup:** This group consistently shows the lowest symptom rates among the compared groups, with a maximum rate of 35.27% and a remarkably low minimum of 10.67%. The average rate stands at 22.74%, reflecting cultural, social, or possibly systemic differences in how mental health issues are reported, treated, or perceived within this community.

4. **Other (Non-Hispanic, Other Races and Multiple Races):** This category presents a unique trend line with variable rates, which could be indicative of the heterogeneous nature of this group. The highest recorded was 46.07%, and the lowest was 24.23%, with an average of 37.89%, suggesting significant variability possibly stemming from the diverse socio-economic and cultural backgrounds represented within this subgroup. 

### Trends in Depressive Disorder Rates by Population Subgroup by Education

This analysis highlights the variations in depressive disorder rates across different educational subgroups in the US, utilizing data from mid 2020 up to early 2024. The visualization provided in the @fig-educandtrend-graph reveals distinct temporal trends across various education demographics, illustrating how depressive symptoms fluctuate over time among these groups.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-educandtrend-graph
#| fig-cap: "Depressive Disorder Trend by Population Subgroup by Education"

library(tidyverse)
library(arrow)
library(here)
library(ggrepel)
library(extrafont)

# Load the fonts
loadfonts(device = "win")

# Define the path to the data file and load it
data <- read_parquet(file = here::here("outputs/data/Educ_subgroup_trends.parquet"))

# Convert 'Date' column to Date format if it's not already
data$Date <- as.Date(data$Date, format = "%Y-%m-%d")

# Rename the subgroup values
data$Subgroup <- recode(data$Subgroup,
                        "Bachelor's degree or higher" = "College graduate",
                        "High school diploma or GED" = "High school graduate",
                        "Some college/Associate's degree" = "Some college",
                        "Less than a high school diploma" = "No high school degree")

# Define the specific dates for the x-axis scale
specific_dates <- as.Date(c("2020-05-05", "2020-09-14", "2021-01-05", 
                            "2021-05-10", "2021-09-13", "2022-01-10", 
                            "2022-05-09", "2022-09-13", "2023-01-16",
                            "2023-05-08", "2023-09-04", "2024-01-08"))

# Prepare data for labels
label_data <- data %>%
  group_by(Subgroup) %>%
  summarize(Date = median(Date), Value = median(Value)) %>%
  ungroup()

# Plotting the line graph with adjusted label positions
p <- ggplot(data, aes(x = Date, y = Value, group = Subgroup, color = Subgroup)) +
  geom_line() +
  geom_text_repel(data = label_data, aes(label = Subgroup, x = Date, y = Value), 
                  family = "Times", 
                  size = 3,  # Smaller text size
                  arrow = arrow(type = "closed", length = unit(0.02, "inches"), angle = 30),
                  box.padding = unit(0.5, "lines"),
                  point.padding = unit(0.5, "lines")) +
  geom_hline(yintercept = seq(0, 40, by = 5), color = "lightgrey") +
  scale_x_date(breaks = specific_dates, date_labels = "%Y-%m") +
  labs(x = "Date", y = "Symptom Rates (%)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, family = "Times"),
        axis.text.y = element_text(family = "Times"),
        axis.title.x = element_text(family = "Times"),
        axis.title.y = element_text(family = "Times"),
        legend.position = "none",
        plot.margin = unit(c(1, 1, 1, 1), "cm"))

print(p)

```

@fig-educandtrend-graph presents the depressive disorder rates for several educational subgroups, including those with a college degree or higher, some college or an associate's degree, a high school diploma or GED, and less than a high school diploma. Each line represents the trend within a specific educational subgroup, offering a visual representation of how mental health issues vary based on educational attainment.

1. **Less Than a High School Diploma:** Individuals with less than an high school diploma experienced the highest depressive disorder rates throughout the period, with peaks reaching nearly 47%. 

2. **Some College/Associate's Degree:** Individuals in this group exhibit slightly higher rates than those with higher education but still lower than the less educated groups. This group also showed a maximum rate close to 40%, averaging 26 - 27% rates. 

3. **High School Diploma or GED:** For individuals holding a high school diploma or GED, the depressive rates were slightly lower but still significant, peaking at nearly 39%. It is showing similar pattern with **Some College/Associate's Degree** subgroups. 

4. **College Graduate or Higher:** This subgroup typically shows the lowest rates of depressive symptoms. The Bachelor's degree or higher subgroup exhibited the lowest rates across all metrics, with a peak rate of just over 31% generally averaging 15% during the period. 

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-lonely-age-pie
#| fig-cap: "Adults who usually or always feel lonely by Age"

# Load necessary libraries
library(ggplot2)
library(dplyr)
library(readr)
library(here)

# Load the data
data <- read_parquet(file = here::here("outputs/data/Social_Connection.parquet"))

# Filter data for the specific indicator and group
lonely_data <- data %>%
  filter(Indicator == "Adults who usually or always feel lonely") %>%
  filter(Group == "By Age") %>%
  group_by(Subgroup) %>%
  summarise(Value = sum(Value), .groups = "drop")

# Calculate percentages
lonely_data$Percentage <- lonely_data$Value / sum(lonely_data$Value) * 100

# Create the pie chart
pie_chart <- ggplot(lonely_data, aes(x = "", y = Percentage, fill = Subgroup)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), position = position_stack(vjust = 0.5),
            family = "Times", size = 4) +  # Set text font to Times and adjust size
  labs(title = "Adults who usually or always feel lonely by Age",
       x = NULL, y = NULL, fill = "Age Group") +
  theme_void() +
  theme(legend.position = "right",
        text = element_text(family = "Times"),  # Apply Times font to all text elements
        plot.title = element_text(size = 14))  # Customize plot title

# Print the pie chart
print(pie_chart)

```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| label: fig-lonely-education-pie
#| fig-cap: "Adults who usually or always feel lonely by Education"

# Change filter to "By Education"
lonely_data_education <- data %>%
  filter(Indicator == "Adults who usually or always feel lonely") %>%
  filter(Group == "By Education") %>%
  group_by(Subgroup) %>%
  summarise(Value = sum(Value), .groups = "drop")

lonely_data_education$Percentage <- lonely_data_education$Value / sum(lonely_data_education$Value) * 100

pie_chart_education <- ggplot(lonely_data_education, aes(x = "", y = Percentage, fill = Subgroup)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), position = position_stack(vjust = 0.5),
            family = "Times", size = 4) +
  labs(title = "Adults who usually or always feel lonely by Education",
       x = NULL, y = NULL, fill = "Education Level") +
  theme_void() +
  theme(text = element_text(family = "Times"),
        plot.title = element_text(size = 14))

print(pie_chart_education)

```


# Model

## Model set-up
\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma)
\end{align}

### Model justification


# Discussion

## Findings

@fig-ageandtrend-graph indicates that while all age groups experienced fluctuations in depressive disorder rates, younger adults consistently reported higher rates. This could be due to a variety of factors including economic instability, career pressures, or social isolation, particularly accentuated by the pandemic. Trends of older adults, age 50 or more, might reflect accumulating stresses or health-related issues as they age, especially during challenging times like the pandemic. 
The peaks and valleys in the @fig-ageandtrend-graph correlate with the unfolding of global events, particularly the COVID-19 pandemic, which has been a significant stressor impacting mental health across all demographics. The data shows how these impacts vary significantly with age, suggesting that age-specific interventions might be necessary to address the mental health needs of different populations effectively. 

@fig-sexandtrend-graph suggests that both subgroups follow fairly similar trends over time, with both lines rising and falling in tandem. However, the female population consistently reports higher symptom rates than the male population. This persistent disparity can be indicative of the different ways in which depressive disorders manifest and are reported across genders, possibly influenced by societal roles, expectations, and differing stressors. Higher trends of female group suggests periods of increased stress or exacerbation of symptoms. The higher overall symptom rates in this subgroup might reflect the broader social, economic, and perhaps biological factors that influence mental health disparities between sexes.


Overall, the trends in depressive disorder rates by race and ethnicity @fig-raceandtrend-graph  underscore disparities in mental health experiences among different racial and ethnic groups in the U.S. Trends of the other group might reflect the diverse socio-economic and cultural backgrounds within this subgroup. These differences likely stem from a complex interplay of socio-economic factors, access to mental health services, cultural perceptions of mental health, and possibly systemic inequalities. The quantitative findings of maximum, minimum, and average rates for each subgroup add depth to our understanding of these disparities, highlighting the specific challenges and variabilities within each community. 

Discussion of Results of the education subgroup

Factors such as lower economic stability, limited access to healthcare, and higher life stressors likely contribute to these higher rates. This group's high average rate suggests a consistent burden of depressive symptoms, potentially exacerbated by factors such as economic insecurity and limited access to mental health resources. 

Those with some college or an Associate's degree also showed high variability in depressive disorder rates indicating that even some post-secondary education does not fully mitigate the risk of higher depressive symptoms. This might reflect a mid-level access to resources and varying levels of stress related to educational and career expectations.

High school diploma or GED, The increased rates could be linked to economic challenges, job market limitations, or the stress associated with lower social mobility. This reflects moderate but considerable mental health challenges within this educational subgroup.

Bachelor's degree or higher subgroup which might indicate better access to resources, stronger economic stability, or higher awareness and utilization of mental health services.

The trends across educational subgroups underscore the strong correlation between educational attainment and mental health outcomes. Higher education often provides better opportunities for employment and income, which are crucial determinants of mental health. Conversely, lower educational attainment is closely associated with various life stressors that exacerbate mental health issues.

## Ethical Implication

The data underscores the need for targeted mental health services and interventions that cater specifically to the unique needs of different age groups. Understanding these trends helps in planning public health strategies and allocating resources more effectively to manage and mitigate depressive disorders among the population.

In conclusion, the depressive disorder rates among different age groups provide crucial insights into the mental health landscape during and following significant global crises. These trends highlight the importance of continuous monitoring and responsive public health strategies to support mental wellness across the lifespan.

The depressive disorder rates among different educational groups emphasize the need for policy interventions that consider the intersection of education, economic stability, and mental health. Tailoring mental health services to meet the needs of people across different educational backgrounds can lead to more effective prevention and treatment strategies, ultimately reducing the overall burden of depressive disorders in society.

## Limitation 

The survey was designed to meet the goal of accurate and timely weekly estimates. Hence, it was conducted by an internet questionnaire, with invitations to participate sent by email and text message (National Center for Health Statistics, 2024). Internet questionnaires face several significant challenges. One of the primary concerns is self-selection bias, where the survey tends to attract respondents who are particularly interested in the topic, which can skew the results and limit their generalizability. Additionally, these surveys may not reach a diverse demographic due to varying levels of internet access across different population segments, thus excluding certain groups, such as older adults or lower-income individuals, and potentially leading to a lack of representativeness. Respondents might also misunderstand or misinterpret questions without the opportunity for real-time clarification, compromising the accuracy of the data collected. Privacy and security concerns are prevalent too, as individuals may be hesitant to share personal or sensitive information online for fear of data breaches or doubts about the survey platform's security measures. Lastly, the validity and honesty of responses are at risk; while the anonymity of online surveys can encourage candidness, it might also lead to less thoughtful or dishonest answers, affecting the reliability of the survey outcomes (Dillman, Smyth, & Christian, 2014; Morling, 2018).

## Future Research
Given the self-reported nature of the data, researchers should be cautious about biases inherent in self-report surveys, such as social desirability bias or recall bias. Additionally, because the dataset represents a specific period (COVID-19 pandemic), care should be taken when generalizing findings to other times or situations. Mitigating these risks involves robust statistical techniques, clear reporting of the survey's scope and limitations, and, where possible, validation with other data sources. Furthermore, beyond studying the impact of COVID-19, the dataset could be used for longitudinal studies on public health trends, economic analysis, and social science research. It can serve as a baseline for future crises management studies, or for comparative studies on how different demographics cope with societal stressors.

\newpage


# References

---
nocite: '@*'
---

